|

### 2021-12-10 06-Chp-IPD-Analysis.Rmd
Update IPD chapter with imputed data

Missing variables were originally intended to be addressed using a previously described method.[@fibrinogenstudiescollaboration2009] Here, the correlation between the fully-adjusted and partially-adjusted estimate in cohorts with the full set of covariates is used to estimate the fully-adjusted effect in those cohorts missing covariates. However, this method requires several large cohorts to contain the full set of covariates, a condition this analysis failed to meet given the very low response rate. As such, two other common approaches were employed.[@fibrinogenstudiescollaboration2009] In the primary analysis, all cohorts were analysed adjusted for the set of common covariates across cohorts (Model 1: age, sex, smoking, alcohol, presence of morbidities). As a sensitivity analysis, cohorts with the full complement of covariates (Whitehall II and EPIC) were analysed using a maximally-adjusted model (Model 2: Model 1, further adjusted for education and BMI). Results between the common-set-adjusted (Model 1) and maximally-adjusted (Model 2) models were then compared.

@@

@@

Missing values in collected variables was common across the cohorts. A matrix of covariates, and whether they were not available or contained missing values for each included cohort, can be seen Appendix \@ref(appendix-ipd-covariate-matrix). The Whitehall II and EPIC cohorts contained data on all but one covariates of interest, while three missing covariates were identified in the CaPS cohort, namely education, BMI and ethnicity. No included cohort provided information on the _Apo_$\mathcal{E}4$ status of participants, and so this variable could not be adjusted for in the analysis.

@@

Estimates from the common-set-adjusted model (Model 1: adjusted for age, sex, smoking, alcohol, presence of vascular comorbidites) were comparable to the fully-adjusted model (Model 2: Model 1 further adjusted for education and BMI) for the effect of lipids on all-cause dementia in cohorts reporting all covariates of interest (Figure \@ref(fig:ipdModelComparison)).

<!----------------------------------------------------------------------------->
(ref:ipdModelComparison-cap) Comparison of results from the common-set-adjusted (Model 1) and fully-adjusted (Model 2) analyses.

@@

<!-- TODO If using fixed effect model, need to udate subgroup labels to reflect this -->

@@

In an attempt to validate my approach, I compared the results from the maximally-adjusted model (Model 2) in this analysis with those reported by Tynkkynen _et al._ for both all-cause dementia and Alzheimer's disease. The results were broadly comparable (Figures \@ref(fig:whitehallComparisonDementia) and \@ref(fig:whitehallComparisonAd)), with the exception of the association of triglycerides with all-cause dementia. Tynkkynen _et al._[@tynkkynen2018] found a protective effect of triglycerides on this outcome (`r estimate(0.69,0.56,0.85,"HR")`) while this analysis found evidence for a harmful effect (`r estimate(1.26,1.12,1.41,"OR")`).

@@

<!----------------------------------------------------------------------------->
(ref:whitehallComparisonAd-cap) Comparison of results across the two analysis of the Whitehall II cohort for the Alzheimer's disease outcome.

(ref:whitehallComparisonAd-scap) Comparison of results across the two analysis of the Whitehall II cohort

```{r whitehallComparisonAd, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:whitehallComparisonAd-cap)', out.width='100%', fig.scap='(ref:whitehallComparisonAd-scap)'}
knitr::include_graphics(file.path("figures/hold.png"))
```
<!----------------------------------------------------------------------------->

@@

Investigation of this discrepancy indicated that it was likely to be due to the Whitehall snapshot that each analysis had access to, as comparison of summary statistics indicated that this analysis had substantially more dementia events (N =287 in this analysis vs. n=114 in Tynkkynen _et al._). The covariates adjusted for in each analysis were similar, though the previous analysis had access to genetic data, allowing it to adjust for _Apo_$\mathcal{E}4$ status. However, given that _Apo_$\mathcal{E}4$ is a risk factor for increased LDL-c rather than triglycerides,[@bennet2007] it seems unlikely that additional adjustment for this variable is responsible for the discrepancy in findings observed.

>>>>

Missing variables were originally intended to be addressed using a previously described method.[@fibrinogenstudiescollaboration2009] Here, the correlation between the fully-adjusted and partially-adjusted estimate in cohorts with the full set of covariates is used to estimate the fully-adjusted effect in those cohorts missing covariates. However, this method requires several large cohorts to contain the full set of covariates, a condition this analysis failed to meet given the very low response rate. As such, two other common approaches were employed.[@fibrinogenstudiescollaboration2009] In the primary analysis, all cohorts were analysed adjusted for the set of common covariates across cohorts (Model 1: age, sex, smoking, alcohol, diabetes). As a sensitivity analysis, cohorts with the full complement of covariates (Whitehall II and EPIC) were analysed using a maximally-adjusted model (Model 2: Model 1, further adjusted for education, ethnicity, and prevalent ischemic heart disease and BMI). Results between the common-set-adjusted (Model 1) and maximally-adjusted (Model 2) models were then compared.

@@

<!-- TODO Need to state that using a fixed effect model for the interaction terms -->

@@

Missing values in collected variables was common across the cohorts. A matrix of covariates, and whether they were not available or contained missing values for each included cohort, can be seen Appendix \@ref(appendix-ipd-covariate-matrix). The Whitehall II and EPIC cohorts contained data on all but one covariates of interest, while three missing covariates were identified in the CaPS cohort, namely education, prevalent ischemic heart disease, BMI and ethnicity. No included cohort provided information on the _Apo_$\mathcal{E}4$ status of participants, and so this variable could not be adjusted for in the analysis.

@@

Estimates from the common-set-adjusted model (Model 1: adjusted for age, sex, smoking, alcohol, and diabetes) were comparable to the fully-adjusted model (Model 2: Model 1 further adjusted for ethnicity, prevalent ischemic heart disease, education and BMI) for the effect of lipids on all-cause dementia in cohorts reporting all covariates of interest (Figure \@ref(fig:ipdModelComparison)).

<!----------------------------------------------------------------------------->
(ref:ipdModelComparison-cap) Comparison of results from the common-set-adjusted (Model 1) and fully-adjusted (Model 2) analyses for the all-cause dementia outcome.

@@

@@

In an attempt to validate my approach, I compared the results from the maximally-adjusted model (Model 2) in this analysis with those reported by Tynkkynen _et al._ for both all-cause dementia and Alzheimer's disease.

The results for the Alzheimer's disease outcome were comparable (Appendix \@ref(appendix-ipd-whitehall)). However, for the vascular dementia outcome (Figure \@ref(fig:whitehallComparisonDementia)), I identified a discrepancy in the association of triglycerides with all-cause dementia estimated by the two analyses. Tynkkynen _et al._[@tynkkynen2018] found a protective effect of triglycerides on this outcome (`r estimate(0.69,0.56,0.85,"HR")`) while this analysis found evidence for a harmful effect (`r estimate(1.26,1.12,1.41,"OR")`).

@@

@@

<!-- TODO Talk about the fact that I used multiple imputation and they did not. Evidenced by wider confidence intervals, indicating a smaller sample size, likely the result of a complete case -->

Investigation revealed several possible reasons for this discrepancy. The first is that it was likely to be due to the Whitehall snapshot that each analysis had access to, as comparison of summary statistics indicated that this analysis had substantially more dementia events (N =287 in this analysis vs. n=114 in Tynkkynen _et al._). Additionally, this analysis uses multiple imputation to address missing data, while the Tynkkynen _et al._ does not mention how missing data was addressed, though the reduced precision in the estimates suggest a complete case analysis. Finally, the covariates adjusted for in each analysis were similar, though the previous analysis had access to genetic data, allowing it to adjust for _Apo_$\mathcal{E}4$ status. However, given that _Apo_$\mathcal{E}4$ is a risk factor for increased LDL-c rather than triglycerides,[@bennet2007] it seems unlikely that additional adjustment for this variable is responsible for the discrepancy in findings observed.
